---
title: "Chess Tournament"
author: "Asher Meyers"
date: "February 16, 2016"
output: html_document
---

Load in the stringr library and the dataset TournamentInfo
```{r}
library(stringr)
TournamentInfo <- read.csv(url("https://raw.githubusercontent.com/AsherMeyers/DATA-607/master/tournamentinfo.txt"),
                           sep = "|", stringsAsFactors=FALSE,
                           col.names = c("Pair", "Name", "Total", "R1", "R2", "R3", "R4", "R5", "R6", "R7", ""), 
                           skip = 1)

TournamentInfo <- TournamentInfo[seq(-2,-194,-3),]
TournamentInfo
View(TournamentInfo)
```


Reorganize the TournamentInfo table so that each player's info falls only on a single line.
```{r}
JoinedRows <- data.frame(TournamentInfo[seq(2,2*194/3,2),],TournamentInfo[seq(3,2*194/3,2),])
```

Pre-Rating Location: Identify the start and end positions of the player's rating in a string, and then identify the content of that substring and save it as the player's pre-tournament rating.
```{r}
PreRatingLoc <- c(str_locate(JoinedRows[,13],"R: ")[1,2]+1, str_locate(JoinedRows[,13], "R: ")[1,2]+4)
PreRatings <- as.integer(str_sub(JoinedRows[,13],PreRatingLoc[1],PreRatingLoc[2]))
```

Write a table with Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre-Chess Rating of Opponent
```{r}
ChessTable <- data.frame(c(1:64), sapply(JoinedRows[2],str_trim),JoinedRows[12],JoinedRows[3], PreRatings)
colnames(ChessTable) <- c("Number", "Name", "State", "Points", "PreRating")
View(ChessTable)
```


Listing of each player's opponents
```{r}
Opponents <- data.frame(1, as.list(as.numeric(str_extract(JoinedRows[1,4:10],"[0-9]{1,2}"))))
colnames(Opponents) <- c("Number", "Opp1","Opp2","Opp3","Opp4","Opp5","Opp6","Opp7")
for (i in 2:64) {
  Opponents <- rbind(Opponents, c(i, as.numeric(str_extract(JoinedRows[i,4:10],"[0-9]{1,2}"))))
}
View(Opponents)
```

Create a listing of each player's opponent's ratings, along with a column for the the number of rounds not played (i.e. column NullCount), and an average rating; for the average, the denominator is the number of games played, which ranges from 2 to 7.

```{r}
OppRatings <- data.frame(c(1:64),ChessTable$PreRating[Opponents$Opp1[1:64]],
                         ChessTable$PreRating[Opponents$Opp2[1:64]],
                         ChessTable$PreRating[Opponents$Opp3[1:64]],
                         ChessTable$PreRating[Opponents$Opp4[1:64]],
                         ChessTable$PreRating[Opponents$Opp5[1:64]],
                         ChessTable$PreRating[Opponents$Opp6[1:64]],
                         ChessTable$PreRating[Opponents$Opp7[1:64]])
colnames(OppRatings) <- c("Player","Opp1R","Opp2R","Opp3R","Opp4R","Opp5R","Opp6R","Opp7R")
OppRatings$Average <- round(rowMeans(OppRatings[2:8],na.rm=TRUE), digits=0)
View(OppRatings)
```

Now we have all the desired components of our table. 

```{r}
ChessTable <- data.frame(c(1:64), sapply(JoinedRows[2],str_trim),JoinedRows[12],JoinedRows[3], PreRatings, OppRatings$Average)
colnames(ChessTable) <- c("Number", "Name", "State", "Points", "PreRating","OpponentAverage")
View(ChessTable)
ChessTable
```

