---
title: "Regular Expressions"
author: "Asher Meyers"
date: "February 13, 2016"
output: html_document
RPubs: http://rpubs.com/ashernm/Week4_Strings
---

3. Let's encode our original string of names

```{r}
library(stringr)

name <- c("Moe Szyslak", "Burns, C. Montgomery", "Rev. Timothy Lovejoy", "Ned Flanders", "Simpson, Homer", "Dr. Julius Hibbert")
name
```

3a. Reorder names that start with their last name.


Split those names separated by a comma into two vectors, trim, and reverse order.
```{r}
splitname <- sapply(sapply(str_split(name, ","), str_trim),rev)
splitname
```

Then reverse the vector order, paste the vectors together with a space in between, trim the unnecessary space - but only for vectors that had their last name first. Then turn the list back into a vector
```{r}

for (i in 1:length(name)) {
  splitname[i]<- paste(unlist(splitname[i]), collapse = " ")
}

splitname <- unlist(splitname)
splitname

```

The code summed together:

```{r}
splitname <- sapply(sapply(str_split(name, ","), str_trim),rev)
for (i in 1:length(name)) {
  splitname[i]<- paste(unlist(splitname[i]), collapse = " ")
}
splitname <- unlist(splitname)
splitname

```

3b. Identify if the names in the vectors have a title - I use the rule "if a name starts with two or more letters followed by a period, it has a title"
```{r}
title <- str_detect(name, "^[[:alpha:]]{2,}\\.")
title
```

3c. We'll remove the titles, and then count the number of spaces in each vector element. If the number of spaces is more than one, we conclude someone has more than one given name.
```{r}
untitled <- str_trim(str_replace(splitname, "^[[:alpha:]]{2,}\\.",""))
polynymous <- str_count(untitled, " ") > 1
polynymous
```


Problem 7. Correct expression of extracting HTML tag
```{r}
breaking <- c("<title>+++BREAKING NEWS+++</title>")
extract <- str_extract(breaking, "<.+>")
```

The regular expression interprets this as "grab every character after  <"
A corrected version follows:

```{r}
breaking <- c("<title>+++BREAKING NEWS+++</title>")
str_extract(breaking, "^<[:alnum:]+>")
```

Problem 8. Extract a formula from a string

```{r}
binomialstring <- c("(5-3)^2=5^2-2*5*3+3^2 conforms to the binomial theorem")
str_extract(binomialstring, "[^0-9=+*()]+")
```

This doesn't work, because the caret and the hyphen are metacharacters that need to be preceded by a double slash (i.e. \\ ), to indicate that those characters are being referenced specifically.

```{r}
str_extract(binomialstring, "[\\^\\-0-9=+*()]+")
```

Alternatively, since the formula is comprised of non-alphabet characters, we can simply exclude the alphabet characters, and trim any hanging spaces.
```{r}
str_trim(str_extract(binomialstring, "[^[:alpha:]]+"))
```

```{r}
code <- c("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")
code <- str_replace_all(code, "e", "*")
alphanum <- c(LETTERS, letters, 0:9,"\\.","\\!")
#str_replace_all(code, alphanum[which.max(charcount)], "e")

```
