---
title: "Project 4"
author: "Asher Meyers"
date: "April 20, 2016"
output: html_document
---

```{r}
install.packages("RMySQL")
library(RMySQL)
```

## R Markdown

First, we access the MySQL file through the package RMySQL, to write the flights to a csv file on a local drive.

```{r}
flightsDB <- dbConnect(MySQL(), user='root', dbname='flights', host='localhost')

airlines <- dbReadTable(flightsDB, "airlines")
airports <- dbReadTable(flightsDB, "airports")
flights <- dbReadTable(flightsDB, "flights")

write.table(airports, file = "C:/flights/airports.csv", sep = ",", row.names = FALSE)
write.table(airlines, file = "C:/flights/airlines.csv", sep = ",", row.names = FALSE)
write.table(flights, file = "C:/flights/flights.csv", sep = ",", row.names = FALSE)

```

To separate these two segments, I re-read the tables from the hard drive, but this is not necessary.


```{r}
flights <- read.csv("C:/flights/flights.csv")
airports <- read.csv("C:/flights/airports.csv")
airlines <- read.csv("C:/flights/airlines.csv")

```



We'll identify what the most frequently occurring airports in our flights arrival locations are. In addition to the three originating airports, namely JFK, LaGuardia and Newark, we will select the 7 most frequent airport destinations. These 7 airports are:

```
 [1] "ORD" "ATL" "LAX" "BOS" "MCO" "CLT" "SFO" 
```
![airports]
(https://raw.githubusercontent.com/AsherMeyers/DATA-607/master/Project4/Figure%201%2010%20Airports.JPG)

Initially, I had selected the top 20, but this made graphical presentations a bit harder, so I opted to simplify. This will also make it easier to manually locate the airports in Neo4j's graphical interface.


```{r}
airportlist <- unique(flights$dest) #List of airports that are destinations for our flights

flightsairports <- table(flights$dest) #frequency counts of flight destinations
flightsairports <- sort(flightsairports, decreasing = TRUE) #sort airports by how many times they appear as destinations
airportlist <- names(flightsairports[1:7])#Pick the 7 most frequently occurring airport destinations
airportlist <- append(airportlist, c("EWR", "JFK", "LGA")) #Adding Newark, JFK and LaGuardia airports

airports3 <- airports[airports$faa %in% airportlist,1:2] #Pull out only the airports that are in the flights selected.

write.table(airports3, file = "C:/flights/airports3.csv", sep = ",", row.names = FALSE)
```


Since there are thousands of flights, and we will be using a graphical interface to understand, we will narrow our flights to a random 50 flights whose destinations are among the 10 airports we've selected. 

For the fields of the airline name, we will selet the plane's number, the flight number, the origin, and the destination.


```{r}
flights10A <- flights[flights$dest %in% airports3$faa, 8:12] #Flights arriving at our 10 airports
flights50 <- flights10A[sample(1:nrow(flights10A),50),] #random sample of 50 flights arriving at our 10 airports
write.table(flights50, file = "C:/flights/flights3.csv", sep = ",", row.names = FALSE)
```

We'll use a list of airlines, but we'll subset that list to only airlines that are flying the flights we selected.

```{r}
airlines3 <- airlines[airlines$carrier %in% flights50$carrier,]
write.table(airlines3, file = "C:/flights/airlines3.csv", sep = ",", row.names = FALSE)
```


```{r}
flightsRoutes <- flights10A[sample(1:nrow(flights10A),100),c(1,4:5)]
airlinesRoutes <- airlines[airlines$carrier %in% flightsRoutes$carrier, 1:2] #Pull out only the airlines that are in the flights selected.
write.table(airlinesRoutes, file = "C:/flights/airlinesRoutes.csv", sep = ",", row.names = FALSE)
write.table(flightsRoutes, file = "C:/flights/flightsRoutes.csv", sep = ",", row.names = FALSE)
```




##Carrier routes

Load airports:

load csv with headers from "file:///flights/airports3.csv" as airports create (a1:Airport {faa: airports.faa, name: airports.name})

Load the airlines:

load csv with headers from "file:///flights/airlinesRoutes.csv" as airlines create (a1:Airline {carrier:airlines.carrier, name:airlines.name})


Create the relationships:

load csv with headers from "file:///flights/flightsroutes.csv" as flights match (a: Airline {carrier:flights.carrier}), (b: Airport {faa:flights.dest}) create (a) -[r:DEST] -> (b)

load csv with headers from "file:///flights/flightsroutes.csv" as flights match (a: Airline {carrier:flights.carrier}), (b: Airport {faa:flights.origin}) create (b) -[r:ORIG] -> (a)




##Let's take a deeper look at one airline's network

First, we'll identify the airline that runs the most flights in our flights dataset.

```{r}
table(flights$carrier) #The number of flights flown by each airline
head(names(sort(table(flights$carrier), decreasing = TRUE)), 1) #The name of the airline with the most flights

```

It happens to be United Airlines, with 58,665 flights out of a total of 336,776, for a share of 17.4%. We'll create a CSV for just United Airlines


```{r}
flightsUA <- subset(flights, flights$carrier == "UA")[,12] #Subset flights into UA flights, carrying only the destination information
airlinesUA <- subset(airlines, airlines$carrier == "UA")
write.table(airlinesUA, file = "C:/flights/airlinesUA.csv", sep = ",", row.names = FALSE)
```

Let's identify the most frequent destinations for this set of flights.

```{r}
flightsUA <- subset(flights, flights$carrier == "UA")[,c(8,12)] #Subset flights into UA flights, carrying only the destination information
Top10NY_UADest <- names(sort(table(flightsUA$dest), decreasing = TRUE)[1:10]) #The top 10 UA destinations for flights leaving NYC
airportsUA <- airports[airports$faa %in% Top10NY_UADest,1:2] #Pull out only the airports that are in the flights selected.

write.table(airportsUA, file = "C:/flights/airportsUA.csv", sep = ",", row.names = FALSE)
```


Those 10 destination airports are:

"ORD" "IAH" "SFO" "LAX" "DEN" "BOS" "MCO" "FLL" "LAS" "TPA"

Let's subset our flights to 500 random UA flights flying to those airports.

```{r}
flightsNY_UA <- flightsUA[flightsUA$dest %in% Top10NY_UADest,]#Subset our UA flights to those destined for the top 10 UA destinations from NYC area airports
flightsUA500 <- flightsNY_UA[sample(1:length(flightsNY_UA$dest), 500),] #500 random UA flights flying out of NYC to one of the top 10 airports
write.table(flightsUA500, file = "C:/flights/flightsUA500.csv", sep = ",", row.names = FALSE)
```

Now, for our neo4j queries to show United Airlines' frequent destination map:

Load the airline:

load csv with headers from "file:///flights/airlinesUA.csv" as airlines create (a1:Airline {carrier:airlines.carrier, name:airlines.name})

Load the destination airports:

load csv with headers from "file:///flights/airportsUA.csv" as airports create (a1:Airport {faa: airports.faa, name: airports.name})




Relate the destinations to the airline

load csv with headers from "file:///flights/flightsUA500.csv" as flights match (a: Airline {carrier: flights.carrier}), (b: Airport {faa: flights.dest}) create (a) - [r:ARRIVES_AT] -> (b)



Create the United Airline node:

Create (a: Airline {faa: 'UA', name: 'United Airlines'})



Now that we have our datasets, let's migrate them to Neo4J. 


###Load the airports

```
load csv with headers from "file:///flights/airports3.csv" as airports create (a1:Airport {faa: airports.faa, name: airports.name, lat: airports.lat, lon:airports.lon, alt:airports.alt, tz:airports.tz, dst:airports.dst})
```


###Load the flights

```
load csv with headers from "file:///flights/flights3.csv" as flights create (a1:Flight {carrier:flights.carrier, tailnum:flights.tailnum, number:flights.flight, origin:flights.origin, dest:flights.dest})
```

###Relate the flight Departures to the airports

```
load csv with headers from "file:///flights/flights3.csv" as flights match (a: Flight {number: flights.flight}), (b: Airport {faa: flights.dest}) create (a) -[r:ARRIVES ]-> (b)
```

###Relate the flight origins to the airports

Relating the origins of flights to the airports

```
load csv with headers from "file:///flights/flights3.csv" as flights match (a: Flight {number: flights.flight}), (b: Airport {faa: flights.origin}) create (a) -[r:DEPARTS ]-> (b)
```

###Load the airlines

load csv with headers from "file:///flights/airlines.csv" as airlines create (a1:Airline {carrier:airlines.carrier, name:airlines.name})



###Relate the flights to their carrier

load csv with headers from "file:///flights/flights3.csv" as flights match (a: Flight {number:flights.flight}), (b: Airline {carrier: flights.carrier}) create (a) -[r:CARRIED_BY] -> (b)


###Load the planes
```
load csv with headers from "file:///flights/planes.csv" as planes create (a1:Plane {tailnum:planes.tailnum, year:planes.year, type:planes.type, manufacturer:planes.manufacturer, model: planes.model, engines:planes.engines, seats:planes.seats, speed:planes.speed, engine:planes.engine})
```



###Relate the flight arrivals to the airports
The command is as follows:

```
load csv with headers from "file:///flights/flights3.csv" as flights match (a: Flight {number: flights.flight}), (b: Airport {faa: flights.origin}) create (a) -[r:DEPARTS ]-> (b)
```

One problem however: our flights don't have an origin airport that also serves as another flight's arrival location, so no relationships will be formed from this step.


####Relate the airlines to the flights

```
load csv with headers from "file:///flights/flights3.csv" as flights match (a: Flight {number: flights.flight}), (b: Airline {carrier: flights.carrier}) create (a) -[r:FLOWN_BY ]-> (b)
```
