---
title: "FavoriteBooks"
author: "Asher Meyers"
date: "March 19, 2016"
output: html_document
---

First, we will load the packages required to complete the assignment, XML and RJSONIO

```{r}
library("XML")
library("RCurl")
library("plyr")
library("RJSONIO")
library("knitr")
```

On my Github, I have handwritten my choice of three books in three different file formats: HTML, XML and JSON.

The three books are: The Bostonians, Small is Beautiful, and How to Lie with Statistics.

For each book, I gathered the following information and put it into a table by hand:

*Title
*Author(s)
*Publication Date
*Publisher
*Type (i.e. fiction or nonfiction)
*Genre

Let's start with HTML. 

I used this testing page to ensure I had the table I desired in HTML: http://www.w3schools.com/html/tryit.asp?filename=tryhtml_default

My raw HTML code is located here: https://github.com/AsherMeyers/DATA-607/blob/master/Week-8/Books.html

```{r}
#Identify the URL where the HTML table is located
  html.url <- "https://raw.githubusercontent.com/AsherMeyers/DATA-607/master/Week-8/Books.html"

#Download the contents of that HTML
books.html <- getURL(html.url)

#Read the HTML into a table in R
books.html.table <- readHTMLTable(books.html, header = TRUE)

View(books.html.table)
kable(books.html.table)
```

Now for reading the table in XML.

```{r}
books.xml.url <- getURL("https://raw.githubusercontent.com/AsherMeyers/DATA-607/master/Week-8/Books.xml", ssl.verifyPeer=FALSE) #RCurl breaks when confronted with SSL verification, so we set the verify peer field to false
books.xml.data <- xmlParse(books.xml.url) #Parses the HTML file into an R structure
books.xml.table <- ldply(xmlToList(books.xml.data), data.frame) #converts each list in the books.xml file into a component of a dataframe
kable(books.xml.table)

```

In JSON:

```{r}
books.json.url <- "https://raw.githubusercontent.com/AsherMeyers/DATA-607/master/Week-8/Books.json"
books.json.table <- fromJSON(books.json.url)
View(books.json.table)
kable(books.json.table)

```



We see that when importing the table through HMTL, each book gets its own row, while in JSON, each book gets its own column, and the categories are rows; additionally, the headers are rewritten for each book. In XML, the existence of two authors leads to two entries being created for the same book. 

I consulted with my past project partner Chris Martin for how to import files into R for XML.
