---
title: "Week 6 Assignment"
author: "Asher Meyers"
date: "March 5, 2016"
output: html_document
---

First, let's install tidyr and load its library, as well as knitr.
> install.packages("tidyr")

```{r}
library(knitr)
library(tidyr)
```

Let's load the data

```{r}
untidy <- read.csv(url("https://raw.githubusercontent.com/AsherMeyers/DATA-607/master/Week6CSV.csv"), stringsAsFactors = FALSE)
names(untidy) <- c("Airline", "Status", "LA", "Phoenix", "SD", "SF", "Seattle")
kable(untidy)
```

Let's remove the third, blank row and then transpose the data so it's easier to work with.
```{r}
tuntidy <- t(untidy[-3,])
kable(tuntidy)
View(tuntidy)
```


Let's separate the dataframe into two dataframes, one for each airline.
```{r}
cities <- c("LA", "Phoenix", "SD", "SF", "Seattle")
```


```{r}
Alaska <- data.frame(city = cities, airline = tuntidy[1,1], ontime = as.integer(tuntidy[3:7,1]), delayed = as.integer(tuntidy[3:7,2]), stringsAsFactors = FALSE)
Alaska$total <- rowSums(subset(Alaska, select=3:4))
Alaska <- transform(Alaska, AlaskaDelayed = delayed / total)
View(Alaska)
```

```{r}
West <- data.frame(city = cities, airline = tuntidy[1,3], ontime = as.integer(tuntidy[3:7,3]), delayed = as.integer(tuntidy[3:7,4]), stringsAsFactors = FALSE)
West$total <- rowSums(subset(West, select=3:4))
West <- transform(West, WestDelayed = delayed / total)
View(West)
```

Now let's combine and compare how often each airline is delayed:
```{r}
DelayComparison <- data.frame(city = cities,  AlaskaDelayed = Alaska$ProportionDelayed, WestDelayed = West$ProportionDelayed)
DelayComparison <- transform(DelayComparison, DelayPercentDiff = AlaskaDelayed - WestDelayed)
View(DelayComparison)
kable(DelayComparison)
```

We see that Alaska Airlines is uniformly less prone to delay than American West Airlines.
